Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-28 doc/naaclhlt2018-latex/paper.tex
typo

Educational software that aims to teach people new skills, languages, and
academic subjects have become increasingly popular.  The wide-spread
deployment of these tools has created interesting opportunities to study
the process of language acquisition in extremely large samples of learners in naturalistic
situations. The Duolingo shared task on Second Lanugage Acquisition Modeling (SLAM)
was a competitive modeling challenge run in early 2018~\cite{slam18}.
The challenge, organized by Duolingo\footnote{http://duolingo.com}, a popular second
language learning app, was
to use log data from thousands of users completing millions of exercises to 
predict patterns of future translation mistakes in held-out data.  The data set was
divided into three sets covering Spanish speakers learning English ({\tt en\_es}),
English speakers learning Spanish ({\tt es\_en}), and English speakers learning
French ({\tt fr\_en}).
This paper reports the approach used by our team which ended in third place 
for the {\tt en\_es} data set, second place for {\tt es\_en}, and third place
for {\tt fr\_en}.

@@

Our entry to SLAM borrowed conceptually from the Khajah et al. (2016) approach.
Specifically we started with a relatively well known and powerful classification algorithm~\cite[gradient
boosting decision trees, GBDT,][]{ke2017lightgbm}.  This algorithm has some advantages
over deep learning approach including the fact that the \hl{inferred decision rules are
interpretable after the model is fit} while still being able to model relatively complex
decision criteria.  We then created a number of new features for the SLAM
the dataset covering aspect such as user perseverance, learning processes, 
contextual factors, and cognate similarity.  After finding a model which provided the best
held-out performance on the test data set, we conducted a number of ``lesioning" studies where
we selectively removed features from the model and re-estimated the parameters in order
to assess the contribution of particular types of features.  We being by describing our 
overall modeling approach, and then discuss some fo the lessons learned from our analysis.

>>>>

Educational software that aims to teach people new skills, languages, and
academic subjects have become increasingly popular.  The wide-spread
deployment of these tools has created interesting opportunities to study
the process of language acquisition in extremely large samples of learners in naturalistic
situations. The Duolingo shared task on Second Lanugage Acquisition Modeling (SLAM)
was a competitive modeling challenge run in early 2018~\cite{slam18}.
The challenge, organized by Duolingo\footnote{http://duolingo.com}, a popular second
language learning app, was
to use log data from thousands of users completing millions of exercises to 
predict patterns of future translation mistakes in held-out data.  The data was
divided into three sets covering Spanish speakers learning English ({\tt en\_es}),
English speakers learning Spanish ({\tt es\_en}), and English speakers learning
French ({\tt fr\_en}).
This paper reports the approach used by our team which ended in third place 
for the {\tt en\_es} data set, second place for {\tt es\_en}, and third place
for {\tt fr\_en}.

@@

Our entry to SLAM borrowed conceptually from the Khajah et al. (2016) approach.
Specifically we started with a relatively well known and powerful classification algorithm~\cite[gradient
boosting decision trees, GBDT,][]{ke2017lightgbm}.  This algorithm has some advantages
over deep learning approach including the fact that the \hl{inferred decision rules are
interpretable after the model is fit} while still being able to model relatively complex
decision criteria.  We then created a number of new features for the SLAM
the dataset covering aspect such as user perseverance, learning processes, 
contextual factors, and cognate similarity.  After finding a model which provided the best
held-out performance on the test data set, we conducted a number of ``lesioning" studies where
we selectively removed features from the model and re-estimated the parameters in order
to assess the contribution of particular types of features.  We being by describing our 
overall modeling approach, and then discuss some of the lessons learned from our analysis.
